Trends of SDR in Thessaloniki

### 2023-07-10 Build_150/Article_B150.Rmd
Commit 2023-07-10 21:13

<!-- something about aerosols: bais mail -->
<!-- something about clouds: bais mail -->
<!--
Η εισαγωγή είναι εντάξει. Ίσως θα μπορούσες να βάλεις στη 2η
παράγραφο ακόμη 2-3 σειρές για τις τάσεις του αεροζόλ και των νεφών
από τις εργασίες και να καταλήξεις με κάποια στοιχεία για το αεροζόλ
στη Θεσσαλονίκη από εργασίες δικές μας. Για τα νέφη δεν έχω υπόψη μου
καμία εργασία αλλά ίσως θα μπορούσε να ειπωθεί κάτι χρησιμοποιώντας
παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
-->

The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on it's climate.
Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
Can also affect, solar and agricultural production, and living organisms.
Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
<!-- Trends china -->
On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
<!-- Trends Europe Brazil -->
A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
<!-- Trends Global -->
Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].

@@

<!-- time span -->
The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from 
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS$Date)`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS$Date)`.
<!-- }. -->
Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from 
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
<!-- }. -->
Although, we performed a similar analysis to the DNI as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
We have to note also, that despite the capability of the CSid to use the DNI as an characterization parameter, we haven't use it, to avoid any selection bias due to unequal availability of measurements.

@@

The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ38'$N, $22^\circ57'$E, 80m\ a.s.l.).
The monitoring site is located near the city centre.
During the study period the CM-21 pyranometer has been independently calibrated four times.
<!-- Preparation. -->
There are three distinct steps to the creation of this dataset:
a)\ the acquisition of radiation measurements from the sensors, 
b)\ a radiation data quality check,
c)\ the identification of "clear sky" conditions from the radiometric data, and
d)\ data aggregation and trend analysis.

@@

<!-- a) Acquisition -->
For the acquisition of radiometric data,
the signal of the broadband instruments is sampled with a rate of $1 \text{Hz}$. The mean and its standard deviation are recorded for every minute.
<!-- dark correction -->
The measurements are corrected for the zero offset of the instrument signal.
As reference of the "dark signal" is used the Sun elevation angle bellow $-10^\circ$, for a period of $3 \text{h}$.
<!-- physical quantities -->
The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.

@@

cite iterative process

<!-- CS subjectivity -->
We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.

<!-- basic concepts -->
The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.

@@

As SDR clear sky reference we are using the radiation model by Haurwitz’s, adjusted for our site with the factor  $a = `r signif(alpha,3)`$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process.
The selection of a clear sky reference model, was based on SDR observation from the period 2016 – 2021.
Where, after an iterative optimization of eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with a wide range factors.
We found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has one of the lowest root mean squared error (RMSE), while retains most of our data. 
The tried models are described in more details by @Reno2012 and tested by @Reno2016.
The iterative optimization method, for the selection of the reference is described by @Long2000 and @Reno2016.
The target of the optimization was the minimizing of the function .......

\begin{equation}
f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{meas},CS,i} - a \times \text{SDR}_{\text{ref},CS,i})^2
\end{equation}

@@

Of the total 1-minute measurements
$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ 
is characterized with clear-sky conditon, and the rest 
$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
with cloud-sky conditions.

@@

@@

<!-- aggregation -->
Before further analysis and deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
a) for daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points,
b) accordingly, monthly values are computed by daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
The daily seasonal values were indexed based on the day of year number, and the monthly by the corresponding calendar month.
<!-- Seasonal -->
For the seasons of the year trends, we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.).
<!-- Deseasonalization -->
Finally, for each data set, we remove the natural occurring seasonal variation.
This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
<!-- SZA Deseasonalization -->
In addition to the previous aggregation scheme, each dataset was aggregate in $1^\circ$ SZA bins, separate for cases before and after local noon. 
This gave us an approximation of Sun's SZA contribution to the SDR brightening.

@@

Although, using the cumulative sums plots we can reveal some structure within the trends and some difference between the sky conditions [@Regier2019]. 
On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part in the begin of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
For a more uniform trend, we expect the cumulative sum plot of the anomaly, to have a symmetric 'V' shape.
In our case, for all sky conditions we observe a period where the cumulative remain relative steady.
This indicates, that the anomaly was evenly distributed around the climatological mean.
Another distinct feature of the cumulative plots, is the difference, between the sky conditions, of the start of the sums sudden increase, for the latter part of the period.
This can make us to suspected that clear-Sky conditions SDR and cloud-sky SDR is affected by different mechanisms.

@@

.... cite paper low elevation SDR .....

@@

<!-- 
## Zenith angle dependency and seasons of the year.
-->

## Consistency of the trends

@@

... different in aggregation ...

@@

${MSE}={Cost}=\frac{1}{n}\sum_{i=1}^{n}(A_i - B_i)^2$

>>>>

<!-- something about aerosols: bais mail -->
<!-- something about clouds: bais mail -->
<!--
Η εισαγωγή είναι εντάξει. Ίσως θα μπορούσες να βάλεις στη 2η
παράγραφο ακόμη 2-3 σειρές για τις τάσεις του αεροζόλ και των νεφών
από τις εργασίες και να καταλήξεις με κάποια στοιχεία για το αεροζόλ
στη Θεσσαλονίκη από εργασίες δικές μας. Για τα νέφη δεν έχω υπόψη μου
καμία εργασία αλλά ίσως θα μπορούσε να ειπωθεί κάτι χρησιμοποιώντας
παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
-->
....

The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
Can also affect, solar and agricultural production, and living organisms.
Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
<!-- Trends china -->
On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
<!-- Trends Europe Brazil -->
A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
<!-- Trends Global -->
Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].

@@

<!-- time span -->
The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from 
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS$Date)`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS$Date)`.
<!-- }. -->
Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from 
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
<!-- }. -->
Although, we performed a similar analysis to the DNI as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
We have to note also, that despite the capability of the CSid to use the DNI as a characterization parameter, we haven't used it, to avoid any selection bias due to unequal availability of measurements.

@@

The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ38'$N, $22^\circ57'$E, 80m\ a.s.l.).
The monitoring site is located near the city centre.
During the study period the CM-21 pyranometer has been independently calibrated four times.
........ to verify number ........
<!-- Preparation. -->
There are three distinct steps to the creation of this dataset:
a)\ the acquisition of radiation measurements from the sensors, 
b)\ a radiation data quality check,
c)\ the identification of "clear sky" conditions from the radiometric data, and
d)\ data aggregation and trend analysis.

@@

<!-- a) Acquisition -->
For the acquisition of radiometric data,
the signal of the broadband instruments is sampled with a rate of $1 \text{Hz}$. The mean and its standard deviation are recorded for every minute.
<!-- dark correction -->
The measurements were corrected for the zero offset of the instrument signal.
As reference of the "dark signal" is used the Sun elevation angle bellow $-10^\circ$, for a period of $3 \text{h}$.
<!-- physical quantities -->
The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.

@@

<!-- CS subjectivity -->
We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
<!-- basic concepts -->
The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.

@@

As SDR clear sky reference we are using the radiation model by Haurwitz’s, adjusted for our site with the factor  $a = `r signif(alpha,3)`$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process.
The selection of a clear sky reference model, was based on SDR observation from the period 2016 – 2021.
Where, after an iterative optimization of eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with a wide range factors.
We found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has one of the lowest root mean squared error (RMSE), while retains most of our data. 
The tried models are described in more details by @Reno2012 and tested by @Reno2016.
The iterative optimization method, for the selection of the reference is described by @Long2000 and @Reno2016.
The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished by the computing function `stats::optimise` [@RCT2023], which is based on work by @Brent1973.

\begin{equation}
f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{meas},CS,i} - a \times \text{SDR}_{\text{ref},CS,i})^2 \label{eq:minf}
\end{equation}

@@

@@

After the preparation of the data, we have $`r SDR_N_All`$  1-minute measurements, of which,
$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ 
are characterized with clear-sky condition, and the rest 
$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
with cloud-sky conditions.

@@

<!-- aggregation -->
Before further analysis and deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
a) for daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points,
b) accordingly, monthly values are computed by daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
The daily seasonal values were indexed based on the day of year number, and the monthly by the corresponding calendar month.
<!-- Seasonal -->
For the seasons of the year trends, we grouped the mean daily values by season (Winter: December - February, Spring: March - May, etc.).
<!-- Deseasonalization -->
Finally, for each data set, we remove the natural occurring seasonal variation.
This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
<!-- SZA Deseasonalization -->
In addition to the previous aggregation scheme, each dataset was aggregate in $1^\circ$ SZA bins, separate for cases before and after local noon. 
This gave us an approximation of Sun's SZA contribution to the SDR brightening.

@@

Although, using the cumulative sums plots we can reveal some structure within the trends and some difference between the sky conditions [@Regier2019]. 
On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
For a more uniform trend, we expect the cumulative sum plot of the anomaly, to have a symmetric 'V' shape.
In our case, for all sky conditions we observe a period where the cumulative remain relative steady.
This indicates, that the anomaly was evenly distributed around the climatological mean.
Another distinct feature of the cumulative plots, is the difference, between the sky conditions, of the start of the sums sudden increase, for the latter part of the period.
This can make us to suspected that clear-Sky conditions SDR and cloud-sky SDR is affected by different mechanisms.

@@

.... cite paper about low elevation SDR .....

@@

<!--
## Consistency of the trends

@@

... different in aggregation ...
-->

@@
